---
title: "Buffalo Crime + Permits Tracker"
format:
  html:
    title-block-banner: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
source("R/utils.R")
source("R/data_fetch.R")
source("R/data_prep.R")
source("R/maps.R")

site_data <- build_site_data()
window <- site_data$window
generated_label <- format(
  lubridate::with_tz(window$generated_at, tzone = site_config$timezone),
  "%b %d, %Y %I:%M %p %Z"
)
window_label <- sprintf(
  "%s to %s",
  format_date_label(window$start_date),
  format_date_label(window$end_date)
)

crime_neighborhood_rank <- site_data$crime_neighborhoods |>
  sf::st_drop_geometry() |>
  dplyr::arrange(dplyr::desc(crime_count), neighborhood) |>
  dplyr::slice_head(n = 10) |>
  dplyr::transmute(
    Neighborhood = neighborhood,
    `Crime incidents` = format_number_label(crime_count)
  )

permit_neighborhood_rank <- site_data$permit_neighborhoods |>
  sf::st_drop_geometry() |>
  dplyr::arrange(dplyr::desc(value_sum), neighborhood) |>
  dplyr::slice_head(n = 10) |>
  dplyr::transmute(
    Neighborhood = neighborhood,
    `Declared permit value` = format_currency_label(value_sum),
    Permits = format_number_label(permit_count)
  )
```

:::::::::::::::::::::::::::::::::::::::: page-shell
:::::::::::: masthead
::: section-kicker
Buffalo open data
:::

# Crime incidents and permits, refreshed daily.

::: masthead-copy
A simple 30-day snapshot of public safety activity and permit issuance across Buffalo, combining live incident points, permit points, and neighborhood-level context in one place.
:::

::::::::: meta-row
::::: meta-item
::: meta-term
Window
:::

::: meta-detail
`r window_label`
:::
:::::

::::: meta-item
::: meta-term
Updated
:::

::: meta-detail
`r generated_label`
:::
:::::
:::::::::
::::::::::::

::::: section-block
:::: section-heading
::: section-kicker
Summary
:::

## Key signals from the current 30-day window
::::

```{r}
build_kpi_cards(site_data$kpis)
```
:::::

:::::: section-block
:::: section-heading
::: section-kicker
Main map
:::

## Incidents, permits, and demolition activity in one view

Crime incidents are shown first so the map stays readable at city scale. Turn on permit, demolition, or neighborhood layers from the control when you want to compare them.
::::

::: main-map-frame
```{r}
make_main_map(site_data)
```
:::
::::::

:::::: section-block
:::: section-heading
::: section-kicker
Neighborhoods
:::

## Where recent crime counts are concentrated

This supporting map keeps the neighborhood view simple and focused instead of splitting attention across multiple smaller maps.
::::

::: support-map-frame
```{r}
make_summary_map(
  site_data$crime_neighborhoods,
  value_col = "crime_count",
  title = "30-day crime count",
  palette = site_config$palette$crime_fill,
  value_format = format_number_label
)
```
:::
::::::

:::::::::::: section-block
:::: section-heading
::: section-kicker
Rankings
:::

## Neighborhoods and categories worth a closer look
::::

::::::::: table-stack
::: table-card
### Neighborhoods with the most crime incidents

```{r}
knitr::kable(crime_neighborhood_rank, align = c("l", "r"))
```
:::

::: table-card
### Neighborhoods with the highest declared permit value

```{r}
knitr::kable(permit_neighborhood_rank, align = c("l", "r", "r"))
```
:::

:::::: table-card
### Most common crime and permit categories

::::: table-pair
::: mini-table
#### Crime categories

```{r}
knitr::kable(
  site_data$top_crime_types |>
    dplyr::mutate(n = format_number_label(n)),
  col.names = c("Category", "Incidents"),
  align = c("l", "r")
)
```
:::

::: mini-table
#### Permit types

```{r}
knitr::kable(
  site_data$top_permit_types |>
    dplyr::mutate(n = format_number_label(n)),
  col.names = c("Type", "Permits"),
  align = c("l", "r")
)
```
:::
:::::
::::::
:::::::::
::::::::::::

:::::::: {.section-block .notes-block}
:::: section-heading
::: section-kicker
Method
:::

## Notes and sources
::::

::::: notes-grid
::: note-card
### Sources

Data is pulled from Buffalo Open Data during daily render:

-   Crime Incidents: <https://data.buffalony.gov/d/d6g9-xbgu>
-   Permits: <https://data.buffalony.gov/d/9p2d-f3yt>
-   Neighborhoods: <https://data.buffalony.gov/d/ekfg-mtu8>
:::

::: note-card
### Caveats

The reporting window is calculated in `America/New_York` and includes today plus the previous 29 days. Crime data is preliminary, and both tables depend on upstream schema stability from Buffalo Open Data.
:::
:::::
::::::::
::::::::::::::::::::::::::::::::::::::::
